---
title: "lab_07"
author: "sean mussenden"
date: "8/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings
**Task**: Load rvest, janitor and the tidyverse
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse,rvest,janitor
library(rvest)
library(tidyverse)
library(janitor)
```


Q1. How many individual cases has the U.S. Department of Justice filed against people accused of fraud related to the PPP or EIDL loan program, according to the DOJ website listing those cases: https://www.justice.gov/criminal-fraud/cares-act-fraud?  An example of one case that appears on this page is "U.S. v. Richard Ayvazyan et al". To answer this question, you will need to use rvest to scrape the data on this single webpage into a dataframe that lists only case names. Hint: you will need to make use of html_elements() and html_text() -- a function that extracts text inside of an html tag -- for this.
A1.
There are 101 individual cases.
```{r}
fraud <-"https://www.justice.gov/criminal-fraud/cares-act-fraud"
fraud_html <-fraud %>% 
read_html() %>%
  html_elements('li b') %>% 
  html_text()

clean_fraud <-as_tibble(fraud_html)
```

Q2. In how many individual judicial districts has the U.S. Department of Justice filed cases against people accused of fraud related to the PPP or EIDL loan program, according to the DOJ website listing those cases: https://www.justice.gov/criminal-fraud/cares-act-fraud?  Note: an example of a judicial district is "Southern District of Florida". You will need to use rvest scrape the data on this single webpage into a dataframe.
A2.
There are 31 jurisdictions
```{r}
fraud <-"https://www.justice.gov/criminal-fraud/cares-act-fraud"
jurisdictions <- fraud%>% 
  read_html() %>%
  html_elements('p b i') %>% 
  html_text()

jurisdictions <- as_tibble(jurisdictions)
```

Q4. The website (https://www.justice.gov/criminal-fraud/cares-act-fraud) shows that the DOJ filed more cases in the Southern District of Florida than in any other district. One of those cases was filed against someone named "Diamond Blue Smith". Who is Smith, and what was he accused of, according to the criminal complaint? If you were an editor, would you have assigned a reporter to write a story about this case when the complaint was filed in court? Why or why not?
A4.
Smith, 37 years old recording artist, was sentenced to 20 months after pleading guilty to conspiracy to commit wire fraud. He obtained a PPP loan for his company, Blue Star Records LLC, using false information. He used the false PPP money on luxury items, including a Ferrari. 

Yes, I would assign a reporter to cover this story because it is a glaring example of how the system can be manipulated. The public deserves to know that PPP money is being used in a fraudulent way to buy luxury goods, rather than helping businesses in need.

Q5. In what percentage of all judicial districts has the U.S. Department of Justice filed cases cases against people accused of fraud related to the PPP or EIDL loan program, according to the DOJ website listing those cases: https://www.justice.gov/criminal-fraud/cares-act-fraud? In answering this question, you should also produce a list of judicial districts where DOJ has NOT filed a case, according to this site.  Note: to answer this question, you will need to scrape a table of all district courts on this up-to-date Wikipedia page under the heading "Active Courts": https://en.wikipedia.org/wiki/List_of_United_States_district_and_territorial_courts  
A5.
32.97%

```{r}
wiki <-'https://en.wikipedia.org/wiki/List_of_United_States_district_and_territorial_courts'

wiki_html <-wiki %>% 
  read_html() %>%
  html_table()

wiki_html <-wiki_html[[3]]

wiki_html

31/94*100
```
Q6. What might explain why, according to this site, the DOJ has filed PPP-related fraud cases in less than half of the country's judicial districts?
A6. Certain states have larger populations, therefore increasing the likelihood of PPP fraud, at least per capita. The data showed a significantly more cases of fraud in Florida, a populated state where it is likely easier to get away with fraud as compared to a small, mid-west state. 

Q7. Which state had the most approved PPP loans per 100,000 population? [This web page](https://dwillis.github.io/jour472files/ppp_scraping_example/index.html) has links to 52 individual web pages, one for each state (plus Washington, D.C. and Puerto Rico). Each of those web pages contains a one-row html table that has the name of the state, the number of approved loans, and the 2019 population. Here's an example for [Alabama](https://dwillis.github.io/jour472files/ppp_scraping_example/states/alabama.html). You'll need to loop over the individual state urls, scrape each individual page and combine the information on each page into a single dataframe to answer this question. Don't forget to calculation the loans per 100,000.
A7.South Dakota
```{r}
states <-'https://dwillis.github.io/jour472files/ppp_scraping_example/index.html'
state_table <-states %>%
   read_html() %>%
  html_table()

state_table <-state_table[[1]]

statetable <-tibble()
for(row_number in 1:nrow(state_table)) {
  each_row_df <- state_table %>%
    slice(row_number)
      url <- each_row_df$url
      
      singlestateinfo <- url %>% 
        read_html() %>% 
        html_table()
      
 singlestateinfo <- singlestateinfo[[1]]  
        
statetable <-statetable %>% 
  bind_rows(singlestateinfo)
  
  }
statetable <-statetable %>% 
  mutate(loans_per_100k = total_ppp_loans/population*100000) %>% 
  arrange(desc(loans_per_100k))

statetable
```
